---
title: "MSP infraspecies vs diet with coinertia and PCR"
output: html_notebook
---


```{r, message=FALSE, warning=FALSE}

library(ade4)
library(dplyr)
library(ggplot2)
devtools::load_all()

```



- compute JSD distance matrix based on the most relevant taxonomic levels identified by permanova
- create a coinertia analysis based on data having both metagenomes and diet data
- H2/CH4 modeling based on co-inertia Principal component regression
- validation on other samples having only food data or metagenomes data (partial co-inertia analysis)


## load infraspecies data table
```{r, message=FALSE, warning=FALSE}

infraspecies = readr::read_csv2(system.file("infraspecies", "infraspecies_table_full.csv", package="IBSFood"))[,-1]


meat_plant_ratio = readr::read_csv2(system.file("tables", "meat_plant_ratio.csv", package="IBSFood"))[,-1]

```

## compute JSD distance matrix



```{r, message=FALSE, warning=FALSE}

load("MSP_keep_infraspecies.rda")

MSP_infraspecies =
rbind(
  infraspecies %>%
    mutate_at(vars(contains("M", ignore.case = F)),function(x){x/sum(x)}) %>%
    #filter(msp_name %in% species_select) %>%
    filter(msp_name %in% MSP_keep_infraspecies$value) %>%
    select(msp_name_partition,contains("M", ignore.case = F) ) %>%
    dplyr::rename(taxa=msp_name_partition),
  


  infraspecies %>%
    mutate_at(vars(contains("M", ignore.case = F)),function(x){x/sum(x)}) %>%
    #filter(msp_name %in% species_select) %>%
    filter(!(msp_name %in% MSP_keep_infraspecies$value)) %>%
    select(msp_name,contains("M", ignore.case = F) ) %>%
    group_by(msp_name) %>%
    summarize_at(vars(contains("M", ignore.case = F)),function(x){sum(x)}) %>%
    ungroup() %>% dplyr::rename(taxa=msp_name)
  ) 


#MSP_infraspecies = infraspecies 


MSP_infraspecies_jsd = 
  BiotypeR::dist.JSD(MSP_infraspecies %>% 
                       mutate_at(vars(contains("M", ignore.case = F)),function(x){x/sum(x)}) %>% select(contains("M",ignore.case = F)))

save(MSP_infraspecies_jsd, file="MSP_infraspecies_jsd.rda")




```



## partial co-inertia analysis

### classical co-inertia analysis



```{r, message=FALSE, warning=FALSE}

load(system.file("biom", "food_unifrac.dist", package="IBSFood"))


A = merge(MSP_infraspecies_jsd %>% broom::tidy(diagonal=TRUE, upper=TRUE),
food_unifrac %>% broom::tidy(diagonal=TRUE, upper=TRUE) , by=c("item1","item2")) %>%
  reshape2::dcast(item1~item2, value.var = "distance.x") %>% .[,-1] %>%
  as.dist()

B = merge(MSP_infraspecies_jsd %>% broom::tidy(diagonal=TRUE, upper=TRUE),
food_unifrac %>% broom::tidy(diagonal=TRUE, upper=TRUE) , by=c("item1","item2")) %>%
  reshape2::dcast(item1~item2, value.var = "distance.y") %>% .[,-1] %>%
  as.dist()


MSP_infraspecies_pco = ade4::dudi.pco(A, scannf = F, nf=10) #check nf later

food_pco = dudi.pco(B %>% ade4::quasieuclid(), scannf = F, nf=10)


infraspecies_food_coi = coinertia(MSP_infraspecies_pco,food_pco, scannf=F, nf=10)


```



### projections of supplementary individuals onto co-inertia components

```{r, message=FALSE, warning=FALSE}



MSP_infraspecies_pco = dudi.pco(MSP_infraspecies_jsd, scannf = F, nf=length(MSP_infraspecies_jsd %>% attr(.,"Labels"))) # this requires to do a PCO on all samples

food_pco = dudi.pco(food_unifrac %>% ade4::quasieuclid(), scannf = F, nf=length(food_unifrac %>% attr(.,"Labels")))


#infraspecies_food_coi = coinertia(infraspecies_pco,food_pco, scannf=F, nf=10)

samples_intersect = 
intersect(
MSP_infraspecies_jsd %>% attr(.,"Labels"),

food_unifrac %>% attr(.,"Labels"))

MSP_infraspecies_pco_tabsup = MSP_infraspecies_pco$tab[-which(row.names(MSP_infraspecies_pco$tab) %in% samples_intersect),]
MSP_infraspecies_pco$tab = MSP_infraspecies_pco$tab[samples_intersect,]
MSP_infraspecies_pco$li = MSP_infraspecies_pco$li[samples_intersect,]
MSP_infraspecies_pco$l1 = MSP_infraspecies_pco$l1[samples_intersect,]

MSP_infraspecies_pco$lw = 1/length(samples_intersect)

food_pco_tabsup = food_pco$tab[-which(row.names(food_pco$tab) %in% samples_intersect),]
food_pco$tab = food_pco$tab[samples_intersect,]
food_pco$li = food_pco$li[samples_intersect,]
food_pco$l1 = food_pco$l1[samples_intersect,]

food_pco$lw = 1/length(samples_intersect)

#MSP_infraspecies_pco
#food_pco

MSP_infraspecies_food_coi = coinertia(MSP_infraspecies_pco,food_pco, scannf=F, nf=length(samples_intersect)) # do a co-inertia on samples intersection

MSP_infraspecies_food_coi_spesup = ade4::suprow(MSP_infraspecies_food_coi,MSP_infraspecies_pco_tabsup) # project samples with only metagenomes
MSP_infraspecies_food_coi_foodsup = ade4::supcol(MSP_infraspecies_food_coi,t(food_pco_tabsup)) # project samples with only food data

colnames(MSP_infraspecies_food_coi_spesup$lisup) = colnames(MSP_infraspecies_food_coi$lX)
colnames(MSP_infraspecies_food_coi_foodsup$cosup) = colnames(MSP_infraspecies_food_coi$lY)



rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup) # all coordinates for metagenomics samples

rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup) # all coordinates for food data samples

MSP_infraspecies_food_coi$RV

```

Coinertia RV coefficient is `MSP_infraspecies_food_coi$RV`


```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

#check  PCoA component projection onto co-inertia components

merge(dudi.pco(MSP_infraspecies_jsd, scannf = F, nf=10)$li[1:3],
      rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[1:3], by="row.names") %>% .[-1] %>%
  cor(method="spearman")


merge(dudi.pco(food_unifrac %>% ade4::quasieuclid(), scannf = F, nf=10)$li[1:3],
      rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup)[1:3], by="row.names") %>% .[-1] %>%
  cor(method="spearman")
                                                                                                    

```

### check metadata associations


```{r, message=FALSE, warning=FALSE}



rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[1:5] %>%
  merge(.,IBSMicrobiota::IBSData$metadata %>%
          filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  select_if(is.numeric) %>%
  select(-Height,-Weight) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("AxcX1","AxcX2","AxcX3","AxcX4","AxcX5")) %>%
  group_by(variable) %>%
  na.omit() %>%
  
  summarise(A1_cor =  cor(AxcX1, value, method = "spearman")     %>% round(2),
            A1_p   =  cor.test(AxcX1, value, method = "spearman")$p.value %>% round(2),
            
            A2_cor =  cor(AxcX2, value, method = "spearman")     %>% round(2),
            A2_p   =  cor.test(AxcX2, value, method = "spearman")$p.value %>% round(2),
            A3_cor=  cor(AxcX3, value, method = "spearman")     %>% round(2),
            A3_p   =  cor.test(AxcX3, value, method = "spearman")$p.value %>% round(2),
            A4_cor=  cor(AxcX4, value, method = "spearman")     %>% round(2),
            A4_p   =  cor.test(AxcX4, value, method = "spearman")$p.value %>% round(2),
            A5_cor=  cor(AxcX5, value, method = "spearman")     %>% round(2),
            A5_p   =  cor.test(AxcX5, value, method = "spearman")$p.value %>% round(2),
            n=n())




#rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[1:7] %>%
rbind(MSP_infraspecies_food_coi_spesup$lisup)[1:7] %>%
  merge(.,IBSMicrobiota::IBSData$metadata %>%
          filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  select_if(is.numeric) %>%
  select(-Height,-Weight,-IBS_subtypes) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("AxcX1","AxcX2","AxcX3","AxcX4","AxcX5","AxcX6", "AxcX7")) %>%
  dplyr::rename(clinical=variable, clinical_value=value) %>%
  reshape2::melt(id.vars=c("clinical","clinical_value")) %>%
    na.omit() %>%
  group_by(clinical,variable) %>%
  mutate(n=n()) %>%
  group_by(clinical,variable,n) %>%
  do(
    
    with(.,cor.test(.$clinical_value,.$value, method = "spearman") %>% broom::tidy())
    
  ) %>%
  ggplot() + geom_tile(aes(x=variable %>% gsub("AxcX","PC",.),y=clinical,fill=estimate,alpha=p.value<0.05)) + 
  scale_fill_gradient2("Rho") + scale_alpha_discrete(range = c(0.5,1)) +
  xlab("Principal component") + ylab("")


rbind(MSP_infraspecies_food_coi$lX)[1:7] %>%
  merge(.,IBSMicrobiota::IBSData$metadata %>%
          filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  select_if(is.numeric) %>%
  select(-Height,-Weight,-IBS_subtypes) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("AxcX1","AxcX2","AxcX3","AxcX4","AxcX5","AxcX6", "AxcX7")) %>%
  dplyr::rename(clinical=variable, clinical_value=value) %>%
  reshape2::melt(id.vars=c("clinical","clinical_value")) %>%
    na.omit() %>%
  group_by(clinical,variable) %>%
  mutate(n=n()) %>%
  group_by(clinical,variable,n) %>%
  do(
    
    with(.,cor.test(.$clinical_value,.$value, method = "spearman") %>% broom::tidy())
    
  ) %>%
  ggplot() + geom_tile(aes(x=variable %>% gsub("AxcX","PC",.),y=clinical,fill=estimate,alpha=p.value<0.05)) + 
  scale_fill_gradient2("Rho") + scale_alpha_discrete(range = c(0.5,1)) +
  xlab("Principal component") + ylab("")


```



```{r, message=FALSE, warning=FALSE}

#rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup)[1:7] %>%
rbind(MSP_infraspecies_food_coi$lY)[1:7] %>%
  merge(.,IBSMicrobiota::IBSData$metadata %>%
          filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  select_if(is.numeric) %>%
  select(-Height,-Weight,-IBS_subtypes) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("AxcY1","AxcY2","AxcY3","AxcY4","AxcY5","AxcY6","AxcY7")) %>%
  dplyr::rename(clinical=variable, clinical_value=value) %>%
  reshape2::melt(id.vars=c("clinical","clinical_value")) %>%
    na.omit() %>%
  group_by(clinical,variable) %>%
  mutate(n=n()) %>%
  group_by(clinical,variable,n) %>%
  do(
    
    with(.,cor.test(.$clinical_value,.$value, method = "spearman") %>% broom::tidy())
    
  ) %>%
  ggplot() + geom_tile(aes(x=variable %>% gsub("AxcY","PC",.),y=clinical,fill=estimate,alpha=p.value<0.05)) + 
  scale_fill_gradient2("Rho") + scale_alpha_discrete(range = c(0.5,1)) +
  xlab("Principal component") + ylab("")

rbind(MSP_infraspecies_food_coi_foodsup$cosup)[1:7] %>%
  merge(.,IBSMicrobiota::IBSData$metadata %>%
          filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  select_if(is.numeric) %>%
  select(-Height,-Weight,-IBS_subtypes) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("AxcY1","AxcY2","AxcY3","AxcY4","AxcY5","AxcY6","AxcY7")) %>%
  dplyr::rename(clinical=variable, clinical_value=value) %>%
  reshape2::melt(id.vars=c("clinical","clinical_value")) %>%
    na.omit() %>%
  group_by(clinical,variable) %>%
  mutate(n=n()) %>%
  group_by(clinical,variable,n) %>%
  do(
    
    with(.,cor.test(.$clinical_value,.$value, method = "spearman") %>% broom::tidy())
    
  ) %>%
  ggplot() + geom_tile(aes(x=variable %>% gsub("AxcY","PC",.),y=clinical,fill=estimate,alpha=p.value<0.05)) + 
  scale_fill_gradient2("Rho") + scale_alpha_discrete(range = c(0.5,1)) +
  xlab("Principal component") + ylab("")

```



```{r, fig.height=4, fig.width=10, message=FALSE, warning=FALSE}

meat_plant_ratio = read.csv2(file=system.file("tables/meat_plant_ratio.csv", package="IBSFood"), row.names=1)


rbind(
  rbind(
    MSP_infraspecies_food_coi$lX[1:7] %>% tibble::rownames_to_column("Patient_ID") %>% dplyr::rename(A1=AxcX1, A2=AxcX2, A3=AxcX3, A4=AxcX4, A5=AxcX5, A6=AxcX6, A7=AxcX7) %>% mutate(model="train",data="metagenomes")  ,
    MSP_infraspecies_food_coi_spesup$lisup[1:7] %>% tibble::rownames_to_column("Patient_ID") %>% dplyr::rename(A1=AxcX1, A2=AxcX2, A3=AxcX3, A4=AxcX4, A5=AxcX5, A6=AxcX6, A7=AxcX7) %>% mutate(model="validate",data="metagenomes")),
  rbind(
    MSP_infraspecies_food_coi$lY[1:7] %>% tibble::rownames_to_column("Patient_ID") %>% dplyr::rename(A1=AxcY1, A2=AxcY2, A3=AxcY3, A4=AxcY4, A5=AxcY5, A6=AxcY6, A7=AxcY7) %>% mutate(model="train",data="food"),
    MSP_infraspecies_food_coi_foodsup$cosup[1:7] %>% tibble::rownames_to_column("Patient_ID") %>% dplyr::rename(A1=AxcY1, A2=AxcY2, A3=AxcY3, A4=AxcY4, A5=AxcY5, A6=AxcY6, A7=AxcY7) %>% mutate(model="validate",data="food"))
) %>%
  merge(.,study_global_metadata, by="Patient_ID") %>%
  dplyr::rename(IBS.Severity.Score = `IBS Severity Score`) %>%
  #merge(.,IBSMicrobiota::IBSData$metadata %>%
  #        filter(Visit=="V4",Sample_type=="Stool"), by="Patient_ID", all.x=TRUE) %>%
  merge(.,nutriscore_DI, by="Patient_ID", all.x=TRUE) %>%
  merge(.,meat_plant_ratio %>% dplyr::rename(Patient_ID=variable) %>% select(Patient_ID,meat_plant_ratio), by="Patient_ID", all.x=TRUE) %>%
  merge(.,genes_richness_1M %>% tibble::rownames_to_column("Patient_ID") , by="Patient_ID", all.x=TRUE) %>%
  mutate(IBS.Severity.Score =  ifelse(Health=="control", 0, IBS.Severity.Score)) %>%
  #select(which(sapply(.,class)=="numeric"),model,data, IBS.Severity.Score) %>%
  select_if(function(col) is.numeric(col) | all(col == .$model) | all(col == .$data)) %>%
  #select(-Height,-Weight,-IBS_subtypes) %>%
  mutate(H2_CH4 = H2/(CH4+1)) %>%
  reshape2::melt(id.vars=c("A1","A2","A3","A4","A5","A6","A7","model","data")) %>%
  dplyr::rename(clinical=variable, clinical_value=value) %>%
  reshape2::melt(id.vars=c("clinical","clinical_value","model","data")) %>%
  na.omit() %>%
  group_by(clinical,variable, model,data) %>%
  mutate(n=n()) %>%
  group_by(clinical,variable,n, model,data) %>%
  filter(n>10) %>%
  do(
    
    with(.,cor.test(.$clinical_value,.$value, method = "spearman") %>% broom::tidy())
    
  ) %>%
  #ggplot() + geom_bar(aes(x=paste(data,model),y=estimate,fill=estimate), stat="identity", position = "dodge") + 
  #facet_grid(clinical~variable) + scale_fill_gradient2("Rho") + ylim(-0.3,0.3)
  ggplot() + geom_tile(aes(x=paste0(model %>% stringr::str_sub(1,1),"\n",data %>% stringr::str_sub(1,1)),y=clinical,fill=estimate)) + 
  scale_fill_gradient2("Rho", mid = "lightgrey") + scale_alpha_discrete(range = c(0.5,1)) +
  #scale_fill_gradient("Rho",low = "lightblue",  high ="lightred") + scale_alpha_discrete(range = c(0.5,1)) +
  xlab("Co-inertia principal components\n(t=training set, v=validation set,\n f=Food data, m=Metagenome data)") + ylab("") + facet_wrap(~variable, nrow=1) + scale_y_discrete(labels= c("Age","BMI","IBS-SSS",expression(H[2]),expression(CH[4]),"FSA-NPS diet index","Meat/plant ratio","Richness",expression(paste(H[2],"/",CH[4]," ratio")))) -> p_coinertia_metadata


p_coinertia_metadata


```



## co-inertia figure

```{r}

# 
# ggplot() + 
#   geom_segment(data= merge(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi$lY, by="row.names"), aes(x=AxcX1,y=AxcX2,xend=AxcY1,yend=AxcY2),col="grey") +
#   geom_point(data = rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup), aes(AxcX1,AxcX2), col="lightblue") +
#   geom_point(data = rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup), aes(AxcY1,AxcY2), col="pink") +
#   xlab("Co-inertia axis 1") + ylab("Co-inertia axis 2")



round(MSP_infraspecies_food_coi$eig/sum(MSP_infraspecies_food_coi$eig),2)[1:20] %>% 
  cumsum() %>% as.matrix() %>% as.data.frame() %>% tibble::rowid_to_column("PC") %>%
  ggplot() + geom_bar(aes(y=V1,x=PC), stat="identity", width = 0.5) + 
  scale_y_continuous(label=scales::percent, breaks =  scales::pretty_breaks(n = 8)) + 
  scale_x_continuous(breaks =  scales::pretty_breaks(n = 7), limits = c(0,8)) +
  geom_hline(yintercept = 0.5, linetype=2, color="red" ) + ylab("Cumulative variance (%)")



rbind(
  rbind(
    MSP_infraspecies_food_coi$lX[1:2] %>% dplyr::rename(A1=AxcX1, A2=AxcX2) %>% mutate(Set="Training set",Data="Metagenomes"),
    MSP_infraspecies_food_coi_spesup$lisup[1:2] %>% dplyr::rename(A1=AxcX1, A2=AxcX2) %>% mutate(Set="Validation set",Data="Metagenomes")),
  rbind(
    MSP_infraspecies_food_coi$lY[1:2] %>% dplyr::rename(A1=AxcY1, A2=AxcY2) %>% mutate(Set="Training set",Data="Food"),
    MSP_infraspecies_food_coi_foodsup$cosup[1:2] %>% dplyr::rename(A1=AxcY1, A2=AxcY2) %>% mutate(Set="Validation set",Data="Food"))
) %>% 
  ggplot() + 
  geom_segment(data= merge(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi$lY, by="row.names"), aes(x=AxcX1,y=AxcX2,xend=AxcY1,yend=AxcY2),col="grey") +
  geom_point(aes(x=A1,y=A2, shape=Set, color=Data), alpha=0.7, size=2) +
  xlab("Co-inertia axis A1 (12.48%)") + ylab("Co-inertia axis A2 (8.48%)") + scale_color_brewer(type="qual", palette = 1) -> p_partial_co_inertia


p_partial_co_inertia 


```




## principal component regression

The goal is to predict exhaled gas based on diet and microbiota co-inertia component

### microbiota gas prediction
```{r}

#infraspecies_food_coi$lX => microbiota
#infraspecies_food_coi$lY => food

# infraspecies_food_coi_spesup => microbiota sup
#infraspecies_food_coi_foodsup => food sup

#

rbind(cbind(group="train",MSP_infraspecies_food_coi$lX),cbind(group="test",MSP_infraspecies_food_coi_spesup$lisup))[1:8] %>%
  #merge(.,IBSMicrobiota::IBSData$metadata %>%
  #        filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  mutate(H2_CH4 = log2((H2+1)/(CH4+1))) %>%
  
  ggplot() + geom_jitter(aes(x=(H2+1) %>% log10(),y=(CH4+1) %>% log10()))




rbind(cbind(group="train",MSP_infraspecies_food_coi$lX),cbind(group="test",MSP_infraspecies_food_coi_spesup$lisup))[1:8] %>%
  #merge(.,IBSMicrobiota::IBSData$metadata %>%
  #        filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  #select_if(is.numeric) %>%
  mutate(H2_CH4 = log2((H2+1)/(CH4+1))) %>%
  #filter(H2>5|CH4>5) %>%
  select(Row.names,group,AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,H2_CH4) -> microbiota_data_gas

  


microbiota_data_gas %>%
  filter(group=="train") %>%
  select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,H2_CH4) %>%
  #select(AxcX2,AxcX4,AxcX5,H2_CH4) %>%
  with(
    ., lm(H2_CH4~ 1 + ., data=.)
  ) -> microbiota_model

microbiota_model

gas_microbiota_predict = 
  predict(microbiota_model, 
        microbiota_data_gas %>% 
          filter(group=="test") %>%
          select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,H2_CH4) 
          #select(AxcX2,AxcX4,AxcX5,H2_CH4)
         )

# microbiota_H2_CH4_plot = 
# cbind(predicted_gas=gas_microbiota_predict,observed_gas=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% na.omit %>% as.data.frame %>%
#   ggplot(aes(x=observed_gas %>% as.character %>% as.numeric , y=predicted_gas %>% as.character %>% as.numeric )) + 
#   geom_point() + geom_smooth(method="lm") + 
#   ylab("predicted exaled\nH2/CH4 ratio (log2)") + xlab("observed exaled\nH2/CH4 ratio (log2)") + 
#   #scale_x_continuous(trans="log2") + scale_y_continuous(trans="log2") + 
#   annotate("text", x=-2,y=1, label="r=0.53, p<0.001") #+
#   #xlim(-5,3) + ylim(-5,3)
# 
# microbiota_H2_CH4_plot
# 
# ggsave("microbiota_H2_CH4_plot.pdf")
# 
# 
 cbind(predicted_gas=gas_microbiota_predict,observed_gas=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% na.omit %>% as.data.frame %>%
   
   with(.,cor.test(.$predicted_gas, .$observed_gas, method="pearson"))


```

M157	
M162	
M195	
M226	

```{r}



rbind(cbind(group="train",MSP_infraspecies_food_coi$lY),cbind(group="test",MSP_infraspecies_food_coi_foodsup$cosup))[1:10] %>%
  #merge(.,IBSMicrobiota::IBSData$metadata %>%
  #        filter(Visit=="V4",Sample_type=="Stool"), by.x="row.names", by.y="Patient_ID") %>%
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  #select_if(is.numeric) %>%
  mutate(H2_CH4 = log2((H2+1)/(CH4+1))) %>%
  filter(!(Row.names %in% c("M157","M162","M195","M226"))) %>%
  #filter(H2>5|CH4>5) %>%
  select(Row.names,group,AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,H2_CH4) -> food_data_gas




food_data_gas %>%
  filter(group=="train") %>%
  select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,H2_CH4) %>%
  #select(AxcY2,AxcY4,AxcY5,H2_CH4) %>%
  with(
    ., lm(H2_CH4~ 0 + ., data=.)
  ) -> food_model

food_model

microbiota_data_gas %>%
  filter(group=="train") %>%
  select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,H2_CH4) %>%
  #select(AxcX2,AxcX4,AxcX5,H2_CH4) %>%
   with(
    ., lm(H2_CH4~0 + ., data=.)
  ) -> microbiota_model


gas_food_predict = 
  predict(food_model, 
        food_data_gas %>% 
          filter(group=="test") %>%
          select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,H2_CH4) #%>%
          #dplyr::rename(AxcX1=AxcY1,AxcX2=AxcY2,AxcX3=AxcY3,AxcX4=AxcY4,AxcX5=AxcY5) 
          #select(AxcY2,AxcY4,AxcY5,H2_CH4) #%>%
          #dplyr::rename(AxcX2=AxcY2,AxcX4=AxcY4,AxcX5=AxcY5) 
        )

# 
# 
# cbind(predicted_gas=gas_food_predict,observed_gas=food_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% na.omit %>% as.data.frame %>%
#   filter(predicted_gas < 0.25 & observed_gas > -2) %>%
#   ggplot(aes(x=observed_gas %>% as.character %>% as.numeric , y=predicted_gas %>% as.character %>% as.numeric )) + 
#   geom_point() + geom_smooth(method="lm") + 
#   ylab("predicted exaled\nH2/CH4 ratio (log2)") + xlab("observed exaled\nH2/CH4 ratio (log2)") #+ 
#   #scale_x_continuous(trans="log2") + scale_y_continuous(trans="log2") + 
#   #xlim(0,8) + ylim(0,8)
# 
# 
# 
# cbind(predicted_gas=gas_food_predict,observed_gas=food_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% na.omit %>% as.data.frame %>%
#   filter(predicted_gas < 0.25 & observed_gas > -2) %>%
#   with(.,cor.test(.$predicted_gas, .$observed_gas, method="pearson"))
# 


```




```{r}

rbind(  cbind(ID=food_data_gas %>% filter(group=="test") %>% .$Row.names, predicted_gas=gas_food_predict,observed_gas=food_data_gas %>% filter(group=="test") %>% .$H2_CH4 %>% as.numeric) %>% 
    na.omit %>% as.data.frame %>% mutate(`Validation set`="Individual with\nfood data only"),
  cbind(ID=microbiota_data_gas %>% filter(group=="test") %>% .$Row.names, predicted_gas=gas_microbiota_predict,observed_gas=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4 %>% as.numeric) %>% 
    na.omit %>% as.data.frame %>% mutate(`Validation set`="Individual with\nmetagenomes data only")) %>%
  mutate(predicted_gas = predicted_gas %>% as.character %>% as.numeric, observed_gas =  observed_gas %>% as.character %>% as.numeric) %>%
  mutate(diff = abs(predicted_gas - observed_gas)) %>%
  filter(predicted_gas > 0.5 & observed_gas < 0)
  #group_by(`Validation set`) %>%
  #summarise(max_diff=max(diff), q75_diff = quantile(diff, 0.75)) 



p_gas_prediction =
rbind(
  cbind(predicted_gas=gas_food_predict,observed_gas=food_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% 
    na.omit %>% as.data.frame %>% mutate(`Validation set`="Individual with\nfood data only"),
  cbind(predicted_gas=gas_microbiota_predict,observed_gas=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% 
    na.omit %>% as.data.frame %>% mutate(`Validation set`="Individual with\nmetagenomes data only")) %>%
  
ggplot(aes(x=observed_gas %>% as.character %>% as.numeric , y=predicted_gas %>% as.character %>% as.numeric )) + 
  geom_point(aes(col=`Validation set`)) + geom_smooth(method="lm") + 
  ylab("predicted exaled\nH2/CH4 ratio (log2)") + xlab("observed exaled\nH2/CH4 ratio (log2)") +
  #annotate("text", x=-2,y=1, label="r=0.36, p<0.001") + 
  scale_color_brewer("Validation set", type="qual", palette=1) + 
  facet_wrap(~`Validation set`) #-> p_gas_prediction

p_gas_prediction


rbind(cbind(predicted_gas=gas_food_predict,observed_gas=food_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% 
    na.omit %>% as.data.frame %>% mutate(data="food")#,
  #cbind(predicted_gas=gas_microbiota_predict,observed_gas=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% 
    #na.omit %>% as.data.frame %>% mutate(data="metagenomes")
  ) %>%
  with(.,cor.test(.$predicted_gas, .$observed_gas, method="pearson"))


H2_CH4_modeling=
cbind(predicted=gas_microbiota_predict,observed=microbiota_data_gas %>% filter(group=="test") %>% .$H2_CH4) %>% 
    na.omit %>% as.data.frame %>% mutate(data="metagenomes", model="metagenomes", target="H2_CH4")



```




## final plot


```{r, fig.height=9, fig.width=12}

cowplot::plot_grid(
cowplot::plot_grid(p_partial_co_inertia,  p_gas_prediction, labels = c("A","C")), 
cowplot::plot_grid(p_coinertia_metadata, labels = c("B")), 
nrow=2)

ggsave("Figure3_coinertia_gas_predicted.pdf")


```




## contributor exploration

```{r}

sample_id= 
row.names(rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[1:5])

cor(rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[sample_id,1:5],

MSP_infraspecies %>% 
  mutate_at(vars(contains("M", ignore.case = F)),function(x){x/sum(x)}) %>%
  tibble::column_to_rownames("taxa") %>% 
  t %>% as.data.frame %>% .[sample_id,], method="spearman") -> cor_matrix_coi_infraspecies


cor_matrix_coi_infraspecies %>%
  t %>%
  as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  filter(abs(AxcX1)>0.25 | abs(AxcX2)>0.25 | abs(AxcX3)>0.25 |  abs(AxcX4)>0.25 | abs(AxcX5)>0.25) %>%
  tibble::column_to_rownames("taxa") %>%
  as.matrix() %>% #.[,c(2,4)] %>%
  heatmap(col = cm.colors(256), scale = "col", hclustfun = function(d) hclust(d,method="complete"))



#heatmap(cor_matrix_coi_infraspecies %>% t %>% na.omit)


cor(rbind(MSP_infraspecies_food_coi$lX,MSP_infraspecies_food_coi_spesup$lisup)[1:5])


```



```{r}


sample_id= 
row.names(rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup)[1:5])

cor(
rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup)[sample_id,1:5],

food_otu %>%
  prop.table(2) %>% 
  t %>%
  .[sample_id,], method="spearman") -> cor_matrix_coi_food


heatmap(cor_matrix_coi_food %>% t , col = cm.colors(256), scale="col", hclustfun = function(d) hclust(d,method="complete"))


cor(rbind(MSP_infraspecies_food_coi$lY,MSP_infraspecies_food_coi_foodsup$cosup)[sample_id,1:5])


```


```{r, fig.height=15, fig.width=4, message=FALSE, warning=FALSE}

rownames(cor_matrix_coi_food) = paste0("A",1:5)
rownames(cor_matrix_coi_infraspecies) = paste0("A",1:5)



cbind(cor_matrix_coi_food,cor_matrix_coi_infraspecies) %>%
 t %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  filter(abs(A1)>0.25 | abs(A2)>0.25 | abs(A3)>0.25 |  abs(A4)>0.25 | abs(A5)>0.25) %>%
  tibble::column_to_rownames("variable") %>%
  as.matrix() %>%
  #heatmap(col = cm.colors(256), scale = "col", hclustfun = function(d) hclust(d,method="complete"))
  pheatmap::pheatmap(., clustering_distance_rows = "euclidean", 
           clustering_distance_cols = "euclidean", 
           cutree_rows = 10, 
           annotation_col = NULL, 
           annotation_row = NULL, 
           annotation_colors = NULL, 
           clustering_method = "complete",
           fontsize_row = 6
            ) 

```



### meat plant ratio prediction


```{r}


meat_plant_ratio = read.csv2(file=system.file("tables/meat_plant_ratio.csv", package="IBSFood"), row.names=1)


meat_plant_ratio %>% select(variable, meat_plant_ratio) %>% head


```





```{r}





rbind(cbind(group="train",MSP_infraspecies_food_coi$lX),cbind(group="test",MSP_infraspecies_food_coi_spesup$lisup))[1:8] %>%
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  merge(meat_plant_ratio %>% select(variable, meat_plant_ratio), ., by.y="Row.names", by.x = "variable") %>%
  select(variable,group,AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,meat_plant_ratio) -> microbiota_data

  
rbind(cbind(group="train",MSP_infraspecies_food_coi$lY),cbind(group="test",MSP_infraspecies_food_coi_foodsup$cosup))[1:10] %>%
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  merge(meat_plant_ratio %>% select(variable, meat_plant_ratio), ., by.y="Row.names", by.x = "variable") %>%
  select(variable,group,AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,meat_plant_ratio) -> food_data


microbiota_data %>%
  filter(group=="train") %>%
  select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,meat_plant_ratio) %>%
   with(
    ., lm(meat_plant_ratio~0 + ., data=.)
  ) -> microbiota_model

microbiota_model



food_data %>%
  filter(group=="train") %>%
  select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,meat_plant_ratio) %>%
   with(
    ., lm(meat_plant_ratio~ 0 + ., data=.)
  ) -> food_model

food_model


microbiota_predict = 
  predict(microbiota_model, 
        microbiota_data %>% 
          filter(group=="test") %>%
          select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,meat_plant_ratio) 
          
         )

food_predict = 
  predict(microbiota_model, 
        food_data %>% 
          filter(group=="test") %>%
          select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,meat_plant_ratio) %>%
          dplyr::rename(AxcX1=AxcY1,AxcX2=AxcY2,AxcX3=AxcY3,AxcX4=AxcY4,AxcX5=AxcY5) 
          
        )




cbind(predicted=food_predict,observed=food_data %>% filter(group=="test") %>% .$meat_plant_ratio) %>% 
    na.omit %>% as.data.frame %>% mutate(data="food") %>%
  with(.,{print(cor.test(.$predicted, .$observed, method="pearson")); (cor.test(.$predicted, .$observed, method="spearman"))})




meat_plant_ratio_modeling = 

cbind(predicted=food_predict,observed=food_data %>% filter(group=="test") %>% .$meat_plant_ratio) %>% 
    na.omit %>% as.data.frame %>% mutate(data="food", model="metagenomes", target="meat_plant_ratio")




```



### IBS severity prediction


```{r}





rbind(cbind(group="train",MSP_infraspecies_food_coi$lX),cbind(group="test",MSP_infraspecies_food_coi_spesup$lisup))[1:8] %>%
  
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  mutate(SSgroup = SSgroup %>% as.factor %>% as.numeric) %>%
  mutate(`IBS Severity Score` = ifelse(SSgroup=="control",0, `IBS Severity Score` )) %>%
  
  select(Row.names,group,AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,SSgroup) %>% na.omit() -> microbiota_data_richness


  
rbind(cbind(group="train",MSP_infraspecies_food_coi$lY),cbind(group="test",MSP_infraspecies_food_coi_foodsup$cosup))[1:10] %>%
  
  merge(.,study_global_metadata, by.x="row.names", by.y="Patient_ID") %>%
  mutate(SSgroup = SSgroup %>% as.factor %>% as.numeric) %>%
  mutate(`IBS Severity Score` = ifelse(SSgroup=="control",0, `IBS Severity Score` )) %>%
  
  select(Row.names,group,AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,SSgroup) %>% na.omit() -> food_data_richness


microbiota_data_richness %>%
  filter(group=="train") %>%
  select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,SSgroup) %>%
  
   with(
    ., lm(SSgroup~1 + ., data=.)
  ) -> microbiota_model

microbiota_model



food_data_richness %>%
  filter(group=="train") %>%
  select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,SSgroup) %>%
  
  with(
    ., lm(SSgroup~ 1 + ., data=.)
  ) -> food_model

food_model


richness_microbiota_predict = 
  predict(food_model, 
        microbiota_data_richness %>% 
          filter(group=="test") %>%
          select(AxcX1,AxcX2,AxcX3,AxcX4,AxcX5,SSgroup) %>%
          dplyr::rename(AxcY1=AxcX1,AxcY2=AxcX2,AxcY3=AxcX3,AxcY4=AxcX4,AxcY5=AxcX5)
          
         )


richness_food_predict = 
  predict(food_model, 
        food_data_richness %>% 
          filter(group=="test") %>%
          select(AxcY1,AxcY2,AxcY3,AxcY4,AxcY5,SSgroup) #%>%
          #dplyr::rename(AxcX1=AxcY1,AxcX2=AxcY2,AxcX3=AxcY3,AxcX4=AxcY4,AxcX5=AxcY5) 
        )




cbind(predicted=richness_food_predict,observed=food_data_richness %>% filter(group=="test") %>% .$SSgroup) %>% 
    na.omit %>% as.data.frame %>% mutate(data="food") %>%
  with(.,cor.test(.$predicted, .$observed, method="pearson"))


severity_modeling = 
cbind(predicted=richness_food_predict,observed=food_data_richness %>% filter(group=="test") %>% .$SSgroup) %>% 
    na.omit %>% as.data.frame %>% mutate(data="food", model="food", target="severity")





```



```{r, fig.height=4, fig.width=10}

model_prediction_plot =

rbind(H2_CH4_modeling,
meat_plant_ratio_modeling,
severity_modeling) %>%
  mutate(target= target %>% gsub("H2_CH4", "H2/CH4 ratio",.) ) %>%
  mutate(target= target %>% gsub("meat_plant_ratio", "Meat/plant ratio",.) ) %>%
  group_by(target) %>%
  mutate(predicted = scale(predicted)) %>%
  ggplot() + 
  geom_point(aes(x=observed, y=predicted, color=data, shape=model)) + 
  scale_color_discrete("Validation sets", labels=c("Food","Metagenomes")) +
  scale_shape("Training sets", labels=c("Food","Metagenomes")) +
  facet_wrap(~target, nrow = 1, scales = "free") + 
  ylab("predicted (scaled)") + xlab("observed") 

model_prediction_plot + cowplot::theme_cowplot()


```


## final plot 2

```{r, fig.height=7, fig.width=14}


cowplot::plot_grid(
  cowplot::plot_grid(
    p_partial_co_inertia+ cowplot::theme_cowplot(),  
    model_prediction_plot+ cowplot::theme_cowplot(), 
    labels = c("A","C"), rel_widths = c(1,1.5)), 
    cowplot::plot_grid(
      p_coinertia_metadata+ cowplot::theme_cowplot(),
      NULL, labels = c("B",""), rel_widths = c(9,1) ), 
nrow=2, rel_heights = c(1,1.2))

ggsave("Figure3_coinertia_model.pdf")


```





