---
title: "Diet permanova"
author: "Julien Tap"
date: "12 novembre 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: "hide"
    code_download: true
    df_print: "paged"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=FALSE)
```

# Load library
```{r}

library(vegan)
library(dplyr)


```

# Load data
```{r}
load("metadata_et_grp.rda")
load("food_B_dist.rda")

et_grp
```


# Permanova with Adonis

## Food diet with euclidean distance {.tabset  .tabset-pills} 

### select tabs

select tabs to see different results

### classic adonis
```{r}

vegan::adonis(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),] ) 
vegan::adonis(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),] )
vegan::adonis(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),] )

vegan::adonis(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] )

```

### adonis2 sequential
```{r}
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()
vegan::adonis2(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()
vegan::adonis2(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()

vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()

```


### adonis2 marginal
```{r}
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()

vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
```



## Food diet without euclidean distance {.tabset  .tabset-pills} 

### select tabs

select tabs to see different results

### classic adonis
```{r}

vegan::adonis(B  ~ Gender , data=et_grp[attr(B,"Label"),] ) 
vegan::adonis(B ~ Health , data=et_grp[attr(B,"Label"),] ) 
vegan::adonis(B  ~ cazotype , data=et_grp[attr(B,"Label"),] ) 

vegan::adonis(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] ) 



```


### adonis2 sequencial
```{r}

vegan::adonis2(B  ~ Gender , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()
vegan::adonis2(B  ~ Health , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()
vegan::adonis2(B  ~ cazotype , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()

vegan::adonis2(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] ) %>% broom::tidy()



```


### adonis2 margin
```{r}

vegan::adonis2(B  ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B  ~ Health , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B  ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()

vegan::adonis2(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
```


## Food diet with severity



```{r}

vegan::adonis2(B  ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B  ~ Health , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B  ~ SS , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
vegan::adonis2(B  ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()

vegan::adonis2(B  ~ Gender + SS + cazotype , data=et_grp[attr(B,"Label"),], by="margin" ) %>% broom::tidy()
```



## Visualisation


```{r}

B %>% 
  broom::tidy() %>%
  mutate(item1=item1 %>% as.character(), item2=item2%>% as.character()) %>%
  #filter(item1==item2) %>%
  merge(et_grp %>% select(cazotype), by.y="row.names", by.x="item1") %>%
  merge(et_grp %>% select(cazotype), by.y="row.names", by.x="item2") %>%
  filter(cazotype.x==cazotype.y) %>%
  ggplot() + geom_boxplot(aes(y=distance,x=cazotype.x))


# B %>%
#   as.matrix %>%
#   reshape2::melt() %>%
#   filter(Var1!=Var2) %>%
#   merge(et_grp %>% select(cazotype), by.y="row.names", by.x="Var1") %>%
#   merge(et_grp %>% select(cazotype), by.y="row.names", by.x="Var2") %>%
#   filter()
#   ggplot() + geom_boxplot(aes(y=value,x=paste0(cazotype.x,cazotype.y)))
#   
  

```



