---
title: "Diet permanova"
author: "Julien Tap"
date: "12 novembre 2018"
output: 
  danr::html_danr_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: "hide"
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

# Load library
```{r}

library(vegan)
library(dplyr)
library(ggplot2)
library(tidyr)
library(glue)

```

# Load data
```{r}
load("metadata_et_grp.rda")
load("food_B_dist.rda")


```


# Permanova with Adonis

## Food diet with euclidean distance - betadisper
```{r}
set.seed(2109181)
mod <- betadisper(B %>% ade4::quasieuclid(),et_grp[attr(B,"Label"),"Gender"])
permutest(mod, pairwise = TRUE, permutations = 999)
mod <- betadisper(B %>% ade4::quasieuclid(),et_grp[attr(B,"Label"),"Health"])
permutest(mod, pairwise = TRUE, permutations = 999)
mod <- betadisper(B %>% ade4::quasieuclid(),et_grp[attr(B,"Label"),"cazotype"])
permutest(mod, pairwise = TRUE, permutations = 999)


```

## Food diet with euclidean distance - sequential
```{r}
set.seed(2109181)
vegan::adonis(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),] )
vegan::adonis(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),] )
vegan::adonis(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),] )
vegan::adonis2(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),] )
vegan::adonis2(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] )

```

## Food diet with euclidean distance - sequential - stability
```{r}
OneRun <- function() {
  
  veg_G <- vegan::adonis(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),] )
  veg_H <- vegan::adonis(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),] )
  veg_C <- vegan::adonis(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),] )
  
  res <- c(veg_G$aov.tab$`Pr(>F)`[1], veg_H$aov.tab$`Pr(>F)`[1], veg_C$aov.tab$`Pr(>F)`[1])
  names(res) <- c("Gender","Health","cazotype")
  return(res)
  
}

set.seed(2109181)
for (i in 1:100) {
  if (i == 1) {
    res_glob <- OneRun()
  } else {
    res_glob <- rbind(res_glob,OneRun())
  }
}

res_long <- gather(data.frame(res_glob), variable, pvalue, Gender:cazotype, factor_key=TRUE)

g <- ggplot(res_long, aes(x=pvalue, color=variable)) +
  geom_histogram(fill="white", alpha=0.5, position="identity")
g

```

## Food diet with euclidean distance - marginal
```{r}
# set.seed(2109181)
# vegan::adonis(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" )
# vegan::adonis(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),], by="margin" )
# vegan::adonis(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" )
# 
# set.seed(2109181)
# vegan::adonis(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" )

# set.seed(2109181)
# ord <- dbrda(B %>% ade4::quasieuclid() ~ Gender, data=et_grp[attr(B,"Label"),])
# set.seed(2109181)
# anova(ord, by="terms")
# set.seed(2109181)
# anova(ord, by="margin")

set.seed(2109181)
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" )
vegan::adonis2(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),], by="margin" )
vegan::adonis2(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" )

set.seed(2109181)
vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" )

```

## Food diet with euclidean distance - margin - stability
```{r}
OneRun <- function() {
  
  veg_G <- vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" )
  veg_H <- vegan::adonis2(B %>% ade4::quasieuclid() ~ Health , data=et_grp[attr(B,"Label"),], by="margin" )
  veg_C <- vegan::adonis2(B %>% ade4::quasieuclid() ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" )
  
  res <- c(veg_G$`Pr(>F)`[1], veg_H$`Pr(>F)`[1], veg_C$`Pr(>F)`[1])
  names(res) <- c("Gender","Health","cazotype")
  return(res)
  
}

set.seed(2109181)
for (i in 1:100) {
  if (i == 1) {
    res_glob <- OneRun()
  } else {
    res_glob <- rbind(res_glob,OneRun())
  }
}

res_long <- gather(data.frame(res_glob), variable, pvalue, Gender:cazotype, factor_key=TRUE)

g <- ggplot(res_long, aes(x=pvalue, color=variable)) +
  geom_histogram(fill="white", alpha=0.5, position="identity")
g

```


## Food diet with euclidean distance - margin - stability - several factors
```{r}
OneRun <- function() {
  
  veg <- vegan::adonis2(B %>% ade4::quasieuclid() ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" )
  
  res <- veg$`Pr(>F)`[1:3]
  names(res) <- c("Gender","Health","cazotype")
  return(res)
  
}

set.seed(2109181)
for (i in 1:100) {
  if (i == 1) {
    res_glob <- OneRun()
  } else {
    res_glob <- rbind(res_glob,OneRun())
  }
}

res_long <- gather(data.frame(res_glob), variable, pvalue, Gender:cazotype, factor_key=TRUE)

g <- ggplot(res_long, aes(x=pvalue, color=variable)) +
  geom_histogram(fill="white", alpha=0.5, position="identity")
g

```

## Food diet with euclidean distance - PcoA
```{r}

# PcoA
data <- as.matrix(B %>% ade4::quasieuclid())
pcoa <- cmdscale(data, k = ncol(data)-1, eig=T, add=T)
eig <- eigenvals(pcoa)
var_exp <- 100*eig/sum(eig)

# Merge mapping and PcoA results
data_map <- merge(pcoa$points[,1:4],et_grp[,4:6],by=0)
rownames(data_map) <- data_map$Row.names
data_map <- data_map[,-1]

OnePlot <- function(param) {
  
  p <- vector(mode="list",length=2)
  
  # Loop on comp 1-2 or comp 3-4
  for (j in c(1,3)) {
    
    comp <- c(j,j+1)
    
    p[[comp[2]/2]] <- ggplot(data_map, aes_string(x = glue("V{comp[1]}"), y = glue("V{comp[2]}"),
      group=param, color=param), shape=20) +
      scale_x_continuous(name=glue("Axis {comp[1]} [{round(var_exp[comp[1]],2)} %]")) +
      scale_y_continuous(name=glue("Axis {comp[2]} [{round(var_exp[comp[2]],2)} %]")) +
      geom_hline(yintercept=0, linetype="dashed", color="grey", size=0.3) +
      geom_vline(xintercept=0, linetype="dashed", color="grey", size=0.3) +
      theme_bw() +
      ggtitle(glue("Colored by {param}")) +
      theme(plot.title = element_text(hjust = 0.5, face="bold")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      geom_point(size=1.5) +
      stat_ellipse(type="t", level=0.95)

  }
  
  return(p)
  
}

p <- OnePlot("Gender")
print(p[[1]])
print(p[[2]])
p <- OnePlot("Health")
print(p[[1]])
print(p[[2]])
p <- OnePlot("cazotype")
print(p[[1]])
print(p[[2]])

```



## Food diet with non-euclidean distance - sequential
```{r}
set.seed(2109181)
vegan::adonis(B  ~ Gender , data=et_grp[attr(B,"Label"),] )
vegan::adonis(B ~ Health , data=et_grp[attr(B,"Label"),] )
vegan::adonis(B  ~ cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis2(B  ~ Gender , data=et_grp[attr(B,"Label"),] )
vegan::adonis2(B  ~ Health , data=et_grp[attr(B,"Label"),] )
vegan::adonis2(B  ~ cazotype , data=et_grp[attr(B,"Label"),] )

set.seed(2109181)
vegan::adonis2(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),] )

```



## Food diet with non-euclidean distance - margin
```{r}
set.seed(2109181)
vegan::adonis2(B  ~ Gender , data=et_grp[attr(B,"Label"),], by="margin" )
vegan::adonis2(B  ~ Health , data=et_grp[attr(B,"Label"),], by="margin" )
vegan::adonis2(B  ~ cazotype , data=et_grp[attr(B,"Label"),], by="margin" )

set.seed(2109181)
vegan::adonis2(B  ~ Gender + Health + cazotype , data=et_grp[attr(B,"Label"),], by="margin" )

```



## Food diet with non-euclidean distance - PcoA
```{r}

# PcoA
data <- as.matrix(B)
pcoa <- cmdscale(data, k = ncol(data)-1, eig=T, add=T)
eig <- eigenvals(pcoa)
var_exp <- 100*eig/sum(eig)

# Merge mapping and PcoA results
data_map <- merge(pcoa$points[,1:4],et_grp[,4:6],by=0)
rownames(data_map) <- data_map$Row.names
data_map <- data_map[,-1]

OnePlot <- function(param) {
  
  p <- vector(mode="list",length=2)
  
  # Loop on comp 1-2 or comp 3-4
  for (j in c(1,3)) {
    
    comp <- c(j,j+1)
    
    p[[comp[2]/2]] <- ggplot(data_map, aes_string(x = glue("V{comp[1]}"), y = glue("V{comp[2]}"),
      group=param, color=param), shape=20) +
      scale_x_continuous(name=glue("Axis {comp[1]} [{round(var_exp[comp[1]],2)} %]")) +
      scale_y_continuous(name=glue("Axis {comp[2]} [{round(var_exp[comp[2]],2)} %]")) +
      geom_hline(yintercept=0, linetype="dashed", color="grey", size=0.3) +
      geom_vline(xintercept=0, linetype="dashed", color="grey", size=0.3) +
      theme_bw() +
      ggtitle(glue("Colored by {param}")) +
      theme(plot.title = element_text(hjust = 0.5, face="bold")) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      geom_point(size=1.5) +
      stat_ellipse(type="t", level=0.95)

  }
  
  return(p)
  
}

p <- OnePlot("Gender")
print(p[[1]])
print(p[[2]])
p <- OnePlot("Health")
print(p[[1]])
print(p[[2]])
p <- OnePlot("cazotype")
print(p[[1]])
print(p[[2]])

```

